\section{Related Work}

\subsection{Parallel Method}
\begin{frame}
	\frametitle{Parallel Method}
	\begin{itemize}
		\setlength\itemsep{1em}
		\item
			The wavefront computation is {\em not} cache-friendly.
			\begin{itemize}
				\item 
					To address this cache issue, Maleki et al. developed
					a technique to exploit more parallelism in the
					dynamic programming.
			\end{itemize}
		\item
			The alternative to wavefront method is the traditional
			row-by-row approach.
	\end{itemize}
\end{frame}

\subsection{Variable Gapped Longest Common Subsequence}
\begin{frame}
	\frametitle{Suffix Maximum Query in the VGLCS Algorithm}
	\begin{itemize}
		\setlength\itemsep{1em}
		\item
			Peng's sequential VGLCS algorithm uses disjoint sets for
			suffix maximum query. 
		\item
			We instead use {\em sparse table} to support incremental
			suffix/range maximum queries in our VGLCS algorithm.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Blocked Sparse Table}
	\begin{itemize}
		\setlength\itemsep{1em}
		\item
			Fischer proposed blocked sparse table for better performance
			than the unblocked sparse table, and Fischer's algorithm
			builds least ancestor tables for answering range maximum
			query.
		\item
			We instead use a {\em rightmost-pops} encoding for Cartesian
			trees.
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Cache Issues on Blocked Sparse Table}
	\begin{itemize}
		\setlength\itemsep{1em}
		\item
			Demaine proposed {\em cache-aware} operations on Cartesian
			tree.
		\item
			Masud presents a new encoding method that reduces the number
			of instructions.
		\item 
			Our rightmost-pops encoding for Cartesian trees also reduces
			cache misses, and with a much simpler implementation than
			Demaine's encoding.
	\end{itemize}
\end{frame}